\package{}: an \R{} package for implemetation of the ITHIM model

### 2016-09-12 manuscript/manuscript.tex
Edits to manuscript 1

\begin{figure}[t]
    \centerline{\includegraphics[width=0.4\textwidth]{./figures/hhsmap}}\centerline{\includegraphics[width=0.4\textwidth]{./figures/fig2}}
    \caption{Change in total national DALYs for each regional transportation
estimates.
}\label{dalyFigure}
\end{figure}

@@

%~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
%~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
%~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
%\bibliographystyle{achemnat}
%\bibliographystyle{plainnat}
%\bibliographystyle{abbrv}
%\bibliographystyle{bioinformatics}
%
%\bibliographystyle{plain}
%
% RNA-seq is a powerful technology used to measure genome-wide
% transcription and dozens of statistical methods have been developed
% recently to facilitate its use, including methods for estimating
% expression and identifying differential expression (DE)
% \cite{Oshlack2010}. The ability to evaluate these methods and compare
% their performance in various settings is essential to ensure that
% reliable information is obtained. The Microarray Quality Control
% (MAQC) \cite{shi2006} data sets have proven useful toward this
% end. However, they are limited by small sample sizes and old
% protocol. Evaluation of DE analysis methods can be performed with
% simulated data, and a handful of methods to simulate RNA-seq reads
% have been developed including FluxSimulator and BEERS
% \cite{Griebel2012, Grant2011}.  BEERS is slower than \package{} and,
% as with FluxSimulator, does not allow specification of read count
% means and are therefore not useful for DE-based read simulation.

% We have developed \package{} to simulate reads efficiently under a
% realistic model of DE across two treatment groups, and to use these
% reads to estimate expression levels for downstream analyses, all
% within the \R{} environment. \package{} is built on the software
% package \rsem{} used primarily for transcript quantification
% \cite{li2011}.  To induce differential expression of transcripts we
% adjust the relative transcript abundance across treatment groups in
% the empirical model used by the \rsem{} simulator.  The reads will be
% simulated under the same design as the sample data.  If the sample
% data are paired-end then the simulate reads will also be paired-end.

% Transcript abundances vary substantially across \rnaseq{} experiments.
% Factors such as cell type, organism, and experimental protocol affect
% the expected transcript abundances.  \package{} builds its model with
% sample data from an \rnaseq{} experiment, and uses the empirical read
% count distributions to select a realistic distribution for mean and
% dispersion parameters.  The total expression and fold change
% relationship for the simulated data is also modeled from these data
% and fold change levels are selected to mimic this
% relationship. \package{} also takes advantage of the transcript
% quantification provided by \rsem{} by allowing the user to estimate
% read counts from the simulated reads.  Both the read-based simulation
% and differential expression analysis may be performed within \R{}
% using any of the many \R{} packages available for DE analysis, e.g.,
% DESeq2, EdgeR, EBSeq \cite{love2014, Robinson2010, leng2013}.
% \package{} implements a two stage simulation.  First, it simulates
% read counts per transcript under a negative binomial distribution
% using empirical mean and dispersion estimates.  Second, it calculates
% relative transcript abundance using estimates of transcript length and
% read-count, updates the sample model, and executes the \rsem{} read
% simulator.

% In the first step, read counts for the untreated group are drawn from
% a negative binomial distribution with mean and dispersion parameters
% $\hat{\mu_0}$, $\hat{\varphi_0}$ estimated from the untreated
% group. In the treated group, differentially expressed transcripts are
% simulated using $\mu_1 = \gamma\mu_0$, $\varphi_1=\varphi_0$.  The
% parameter $\gamma$ is chosen to mimic the fold-change vs.\ total
% expression relationship observed empirically, see Figure
% \ref{ma-transcript}.  \sgy{Observed fold-change values decrease with
% expression level.  If fold-change was assigned independently of
% expression level then transcripts with high levels of expression would
% be assigned\ldots}

% Fold change among DE transcripts is chosen by binning the transcripts
% by overall expression into equally sized bins, and using the bin-wise
% standard deviation of the log fold-change to construct an appropriate
% interval for the fold-change, from which a value is sampled randomly.
% The interval is defined as $I_{\textrm{bin}} =
% (k_1\sigma_{\textrm{bin}}, k_2\sigma_{\textrm{bin}})$, where
% $\sigma_{\textrm{bin}}$ is the bin-wise standard deviation of log
% fold-change, and $k_1$ and $k_2$ are chosen by the user.  In the
% following chromosome 22 example, values of $k_1=?$ and $k_2=?$ were
% used with $B=?$ bins.  By tuning parameters $k_1$ and $k_2$ the user
% can control the signal strength while preserving the relationship
% between fold-change and total expression. \sgy{\package{} could
%   easily be adapted to allow user-defined values for $\gamma$, $\mu_i$
%   and $\varphi_i$.}

% Three files are needed for \package{}; a transcript template, the
% read-count estimates for all samples, and a sample model file.  All of
% these are easily created using \rsem{} utilities. A GTF file,
% transcript-to-gene file such as ``knownIsoforms.txt'' from the UCSC
% Genome Browser, and a FASTA reference sequence, or a Multi-FASTA file
% containing transcript sequences, may be used to create the reference
% files.

% In the second stage, the simulated read counts are used to compute the
% relative transcript abundance needed for the read simulation.
% \package{} can return either the gene or isoform level estimated read
% counts from \rsem{}, or simply write the fastq read file.  The
% software package \bowtie{} is used for the alignment
% \cite{langmead2009}.

% We simulated 1,000,000 reads for each of ten samples from two
% treatment groups, and performed DE analysis on all transcripts from
% chromosome 22 using \ebseq{} \cite{leng2013}.  One hundred genes (288
% transcripts) were selected at random to be DE.  In the top of Figure
% \ref{ma-transcript} the log fold-change is plotted versus total
% expression given the simulated counts.  Beneath it is the same figure,
% bottom of Figure \ref{ma-transcript}, but using the estimate counts of
% the simulated reads.  The reads are simulated in such a way as to
% preserve the relationship between fold-change and total expression.

% Using the transcript-wise estimated read counts from the simulated
% data, we performed DE analysis using \ebseq{}.  Of the 288
% differentially expressed transcripts, 179 (62\%) were correctly
% identified by \ebseq{}, with 29 false discoveries.

% %% In this example several of the DE transcripts were assigned a
% %% posterior probability of DE to be zero.  This is due to one of the
% %% treatment groups having zero count.  Also, some of the DE genes are
% %% removed by \ebseq{} because the read count is too low.  Thus some of
% %% the DE transcripts are unidentifiable.  If we remove them we have a
% %% true positive rate of $x$\%.

% Unaligned read data from twenty RNA-seq chips, along with a set of
% chromosome 22 reference files were used to simulate the reads in this
% example.  The example takes about 3 hours to run and may be found in
% the compendium package \compendium{}, available at
% \url{http://www.biostat.wisc.edu/~syounkin/}.

>>>>

\begin{figure}[t]
    \centerline{\includegraphics[width=0.4\textwidth]{./figures/hhsmap}}
    \centerline{\includegraphics[width=0.4\textwidth]{./figures/fig2}}
    \caption{Change in total national DALYs for each regional
      transportation estimates.}\label{dalyFigure}
\end{figure}

@@
